\Large \bf Compilers Project: PTX Super-optimization

### 2018-04-30 paper/main.tex
paper corrections

% In addition, by finding windows with tighter pre and post-conditions,
% the super-optimiser hopes to find better code segments
% that work in the context of the current program.
% This would be very useful for CUDA programs given
% the conditional guards built into PTX: the preconditions
% might be able to be inferred from guards so that
% the program produced is better in the context
% of the larger code segment.
%
% However, the conditional hazards of CUDA -- that
% PTX instructions can be prevented from running
% at the instruction level, means that in order to
% optimise the bulk of the PTX ISA, conditional
% branching cannot be ignored. This will have to be
% a key extension on the original super-optimiser as
% it is a limitation of the current implementation.

@@

Generally, if the environment read in is $(M, I, O, V)$ and
an instruction $o = f(T)$ (where $T$ is some set of inputs)
is read, the environment is updated as follows:
\begin{itemize}
    \item $o \mapsto f(M(T))$ is added to the map (or $o$ is updated).
        $M(T)$ is used in place of $T$ to propagate equations inferred
        as the program is read.
    \item $(T \setminus (V \cup \{o\})) \cup I$
        This removes all variables assigned to before since they
        cannot provide input to the program. Otherwise, unseen variables
        and past inputs are considered inputs.
    \item $(O \setminus T) \cup \{o\}$
        We remove the inputs to this instruction from the output set of
        the program since we assume that all outputs will not be re-used.
        As a workaround, idempotent instructions may be used to produce the
        output that is re-used. The addition of a simple \texttt{add.u32 actual\_out, temporary\_to\_be\_reused, 0;} would suffice.
    \item $V \cup T \cup \{o\}$.
        We maintain this for the second update above: it is indispensable to
        know that variables have actually been seen in order to determine
        which variables are new inputs.
\end{itemize}

@@

The search for \texttt{add.u32 d, a, 0;}
terminates in the idempotent generation phase and is fairly quickly.

@@

\section{Benchmarks and Observations}

@@

@@

>>>>

@@

Generally, if the environment read in is $(M, I, O, V)$ and
an instruction $o = f(T)$ (where $T$ is some set of inputs)
is read, the environment is updated as follows:
\begin{itemize}
    \item $o \mapsto f(M(T))$ is added to the map (or $o$ is updated).
        $M(T)$ is used in place of $T$ to propagate equations inferred
        as the program is read.
    \item $(T \setminus (V \cup \{o\})) \cup I$.
        This removes all variables assigned to before since they
        cannot provide input to the program. Otherwise, unseen variables
        and past inputs are considered inputs.
    \item $(O \setminus T) \cup \{o\}$.
        We remove the inputs to this instruction from the output set of
        the program since we assume that all outputs will not be re-used.
        As a workaround, idempotent instructions may be used to produce the
        output that is re-used. The addition of a simple \texttt{add.u32 actual\_out, temporary\_to\_be\_reused, 0;} would suffice.
    \item $V \cup T \cup \{o\}$.
        We maintain this for the second update above: it is indispensable to
        know that variables have actually been seen in order to determine
        which variables are new inputs.
\end{itemize}

@@

The search for \texttt{add.u32 d, a, 0;}
terminates in the idempotent generation phase and is fairly quick.

@@

\section{Benchmarks, Observations, and Evaluation}

@@

\subsection{Relating to the Problem Definition}

The problem this work intended to solve was to provide a super-optimizer for PTX.

This is certainly a start, but by no means a definitive end. There is much still left to
be learned. In fact, the heuristic currently used is naive and may be futile. The nature
of PTX is not considered: only the length of the assembly. This would certainly correlate
with speed intuitively, but there is no verification of this claim. In fact, confounding
variables such as memory arrangement, the loops, and even instructions themselves are not
considered.

\subsection{Takeaways}

There were two takeaways: that alternatives should be considered sooner and that
the search space always pruned more.

The Ocelot project \cite{GPUOcelot} showed that though a project may seem promising,
its maintenance may not exist. Building the project proved to be impossible as the versions
only aligned when the project was built and that astronomical event has passed.
In the future, less time should be spent behind one project, especially if it does not build.

%Start sooner!
The intractability of the naive search sprang up surprisingly quickly: even generating 2
instruction programs proves to be a 2 minute chore. The output-by-output search definitely
offers a large speed-up. It is interesting that some of the insights that help are rather
simple.

@@

With the above testing, we may be able to verify whether our heuristic of program
length is admissible. This would be another main area of future work where
benchmarks would be invaluable.
